\propertitle

### 2019-12-03 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-03T09:54:02+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

In our case we don't know the time averages of the quantity
$\binom{\yA}{m}$, so we cannot directly equate them to the factorial moment
$\E\bigl[\binom{\yA}{m}\bigr]$ of the larger-population distribution
$\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
step~\ref{item:sample_step} comes to our rescue, because it says that the
expectation for the larger population are determined by that for the sample
$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time average of its
corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
steps:
\begin{equation*}
  \label{eq:chain_prescription_moments}
  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
  \text{measured moments} \rightarrow
        \underbracket{\text{sample moments}
  \rightarrow
    \text{larger-population moments}}_{\textit{sampling theory}}
\end{equation*}
We obtain a distribution $\px(\yA \| M, N)$ for the larger population of
$N$ neurons by constraining some of its factorial moments, for example
$M=\set{1,\dotsc,m'}$ with $m'\le n$, to be equal to the sample's recorded
averages. In formulae, the constraints on $\pxxx(\yA)$ are
\begin{equation}
  \label{eq:final_constraint}
\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
  \equiv
   \sum_{\ya}
  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured averages}}
  \mathrel{\;=\;}
   \underbracket{\sum_{\yA}
     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}},
   \quad m\in M.
\end{equation}

@@

The first and third assumptions above are the least realistic, but they
make sense as reference hypotheses for judging the informational relevance
of correlations of some order, as in \sect~\ref{sec:marginalization}, to be
compared with hypotheses about more realistic time dependences and
clustering.

@@

The space of possible frequency distributions for the larger population has
$N$ dimensions, however, and probability distributions in high dimensions
have counter-intuitive properties. For example, the mode or mean can have
\emph{atypical} features when compared with the features of most other
points of the space. The question, then, is \emph{which features of the
  maximum-entropy frequency distribution are typical of the majority of
  plausible frequency distributions?} We can only answer for sure by using
the full-fledged probability calculus. A more complete study (in
preparation) with the first data set reveals that most of the plausible
frequency distributions have three important features in common with the
maximum-entropy one:
\begin{itemize}
\item all activities $\yA/N \gtrsim 5\%$ have practically zero frequencies;
\item there are two regions of activity levels with high frequencies,
  roughly separated by a trough of lower frequencies.
\item The frequencies of the region on the left ($\yA/N \lesssim 1.8\%$)
  are higher than those of the region on the right ($\yA/N \gtrsim 1.8\%$).
\end{itemize}
But there are also differences. For example, many plausible frequency
distributions have three or four modes instead of just two; these modes are
higher than those of the maximum-entropy distribution; and the bump of high
frequencies on the right is slightly shifted towards lower activities than
the corresponding maximum in the maximum-entropy distribution.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-03T19:25:44+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

In our case we don't know the time averages of the quantity
$\binom{\yA}{m}$, so we cannot directly equate them to the factorial moment
$\E\bigl[\binom{\yA}{m}\bigr]$ of the larger-population distribution
$\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
step~\ref{item:sample_step} comes to our rescue, because it says that the
expectation for the larger population are determined by that for the sample
$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time average of its
corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
steps:
\begin{equation*}
  \label{eq:chain_prescription_moments}
  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
  \text{measured averages} \rightarrow
        \underbracket{\text{sample moments}
  \rightarrow
    \text{larger-population moments}}_{\textit{sampling theory}}
\end{equation*}
We obtain a distribution $\px(\yA \| M, N)$ for the larger population of
$N$ neurons by constraining some of its factorial moments, for example
$M=\set{1,\dotsc,m'}$ with $m'\le n$, to be equal to the sample's recorded
averages. In formulae, the constraints on $\pxxx(\yA)$ are
\begin{equation}
  \label{eq:final_constraint}
\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
  \equiv
   \sum_{\ya}
  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured averages}}
  \mathrel{\;=\;}
   \underbracket{\sum_{\yA}
     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}},
   \quad m\in M.
\end{equation}

@@

The first and third assumptions above are the least realistic, but they
make sense as reference hypotheses for judging the informational relevance
of correlations of some order, as in \sect~\ref{sec:marginalization}, to be
compared with hypotheses about more realistic time dependences and
clustering. They may turn out to approximate well more realistic assumptions.

@@

The space of possible frequency distributions for the larger population has
$N$ dimensions, however, and probability distributions in high dimensions
have counter-intuitive properties. For example, the mode or mean can have
\emph{atypical} features when compared with the majority of other probable
points of the probability space. The question then is \emph{which features
  of the maximum-entropy frequency distribution are typical of the majority
  of plausible frequency distributions?} Preliminary studies using a full
probability calculation (to be published separately) show that several
semi-quantitative features of the maximum-entropy distribution are indeed
typical. For the first data set, for example, most frequency distributions
have two high-frequency regimes of activity, as in the upper panel of
\fig~\ref{fig:plots_5mom}, with roughly the same heights and almost the
same locations, within a few percentages of normalized total activity. Each
of high-frequency regimes, though, may consist of two or three very close
modes rather than one.

This means that the maximum-entropy approximation presented here can be
favourably used to get an idea of what a fully probabilistic calculation
could yield. -- With a far lower computational cost: for example, for
$N=1\,000$ the maximum-entropy distribution can be calculated in ten
minutes on a laptop, whereas the full probability calculation takes several
days on a high-performance computing cluster.
